MS Technical Paper: \\ Placement Algorithms for Heterogeneous FPGAs

### 2025-05-30 latex/paper/sections/09-packing.tex
single site movement
chain swap pseudocode

Packing the other \texttt{Cell} structures into \texttt{Site}s follow a similar process. 
For each cell cluster, we create a \texttt{SiteInst} and populate its \texttt{BEL}s with the corresponding \texttt{Cell}s in the cluster, assign it a physical \texttt{Site} on the \texttt{device}.
We further consolidate these \texttt{SiteInst}s into clusters or lists of \texttt{SiteInst}s for multi-Site structures like \texttt{CarryCellGroup} chains and DSP cascades. 
For each \texttt{SiteInst}, we assign it a physical \texttt{Site} on the \texttt{device} and route its internal intra-Site nets. 
The end result is a \texttt{PackedDesign} object which is a group of data structures solely around \texttt{SiteInst}s like shown in figure \ref{fig:substages}.
At this point we no longer need to worry about individual \texttt{BEL}s, \texttt{Cell}s, or intra-Site \texttt{Nets} as they have now been packaged and taped-off into their respective \texttt{SiteInst}s. 
These \texttt{SiteInst} become the new atomic level components in the following Placement stage.

\end{multicols}
\begin{lstlisting}[caption={\texttt{SiteInst} constructor and methods.}, label={lst:siteinst}]
SiteInst Constructor: 
    SiteInst(String name, Design design, SiteTypeEnum type, Site site)

Most relevant SiteInst Methods:
    createCell(EDIFHierCellInst inst, BEL bel) // Populating the SiteInst BELs with Cells
    unplace() // Unplacing the SiteInst from its current Site
    place(Site site) // Placing an unplaced SiteInst onto a Site
    routeSite() // Attempt to automatically route all intra-Site nets (manual intervention likely required)
    routeIntraSiteNet(Net net, BELPin src, BELPin snk) // Manually route an intra-Site net

@@

\begin{lstlisting}[language=java, caption={Manually reroute intra-Site nets in a \texttt{SLICEL} containing FF \texttt{Cells}}, label={lst:ff_nets}]
private void rerouteFFClkSrCeNets(SiteInst si) {
    si.addSitePIP("CLKINV", "CLK");
    // activate PIPs for SR and CE pins
    Net SRNet = si.getNetFromSiteWire("SRUSEDMUX_OUT");
    Net CENet = si.getNetFromSiteWire("CEUSEDMUX_OUT");
    // if routeSite() default PIPs are incorrect, deactivate them then activate correct PIP
    if (SRNet != null) {
        if (SRNet.isGNDNet()) {
            if (si.getSitePIPStatus(si.getSitePIP("SRUSEDMUX", "IN")) == SitePIPStatus.ON)
                for (String FF : FF_BELS)
                    si.unrouteIntraSiteNet(si.getSitePinInst("SR").getBELPin(), si.getBELPin(FF, "SR"));
            si.addSitePIP(si.getSitePIP("SRUSEDMUX", "0"));
        } else {
            if (si.getSitePIPStatus(si.getSitePIP("SRUSEDMUX", "0")) == SitePIPStatus.ON)
                for (String FF : FF_BELS)
                    si.unrouteIntraSiteNet(si.getBELPin("SRUSEDGND", "0"), si.getBELPin(FF, "SR"));
            si.addSitePIP(si.getSitePIP("SRUSEDMUX", "IN"));
        }
    }
    if (CENet != null) {
        if (CENet.isVCCNet()) {
            if (si.getSitePIPStatus(si.getSitePIP("CEUSEDMUX", "IN")) == SitePIPStatus.ON)
                for (String FF : FF_BELS)
                    si.unrouteIntraSiteNet(si.getSitePinInst("CE").getBELPin(), si.getBELPin(FF, "CE"));
            si.addSitePIP(si.getSitePIP("CEUSEDMUX", "1"));
        } else {
            if (si.getSitePIPStatus(si.getSitePIP("CEUSEDMUX", "1")) == SitePIPStatus.ON)
                for (String FF : FF_BELS)
                    si.unrouteIntraSiteNet(si.getBELPin("CEUSEDGND", "1"), si.getBELPin(FF, "CE"));
            si.addSitePIP(si.getSitePIP("CEUSEDMUX", "IN"));
        }
    }
} // end rerouteFFSrCeNets()
\end{lstlisting}

\begin{multicols}{2}

>>>>

Packing the other \texttt{Cell} structures into \texttt{Site}s follow a similar process. 
For each cell cluster, we create a \texttt{SiteInst} and populate its \texttt{BEL}s with the corresponding \texttt{Cell}s in the cluster and assign it a physical \texttt{Site} on the \texttt{device}.
We further consolidate these \texttt{SiteInst}s into clusters or lists of \texttt{SiteInst}s for multi-Site structures like \texttt{CarryCellGroup} chains and DSP cascades. 
For each \texttt{SiteInst}, we assign it a physical \texttt{Site} on the \texttt{device} and route its internal intra-Site nets. 
The end result is a \texttt{PackedDesign} object which is a group of data structures solely around \texttt{SiteInst}s like shown in figure \ref{fig:substages}.
At this point we no longer need to worry about individual \texttt{BEL}s, \texttt{Cell}s, or intra-Site \texttt{Nets} as they have now been packaged and taped-off into their respective \texttt{SiteInst}s. 
These \texttt{SiteInst} become the new atomic level components in the following Placement stage. 
This process is analogous to the containerization of loose goods in ocean freight forwarding.

\end{multicols}
\begin{lstlisting}[caption={\texttt{SiteInst} constructor and methods.}, label={lst:siteinst}]
SiteInst Constructor: 
    SiteInst(String name, Design design, SiteTypeEnum type, Site site)
Most relevant SiteInst Methods:
    createCell(EDIFHierCellInst inst, BEL bel) // Populating the SiteInst BELs with Cells
    unplace() // Unplacing the SiteInst from its current Site
    place(Site site) // Placing an unplaced SiteInst onto a Site
    routeSite() // Attempt to automatically route all intra-Site nets (manual intervention likely required)
    routeIntraSiteNet(Net net, BELPin src, BELPin snk) // Manually route an intra-Site net

@@

\begin{lstlisting}[language=java, caption={Manually reroute intra-Site nets in a \texttt{SLICEL} containing FF \texttt{Cells}}, label={lst:ff_nets}]
private void rerouteFFClkSrCeNets(SiteInst si) {
    si.addSitePIP("CLKINV", "CLK");
    // activate PIPs for SR and CE pins
    Net SRNet = si.getNetFromSiteWire("SRUSEDMUX_OUT");
    Net CENet = si.getNetFromSiteWire("CEUSEDMUX_OUT");
    // if routeSite() default PIPs are incorrect, deactivate them then activate correct PIP
    if (SRNet != null) {
        if (SRNet.isGNDNet()) {
            if (si.getSitePIPStatus(si.getSitePIP("SRUSEDMUX", "IN")) == SitePIPStatus.ON)
                for (String FF : FF_BELS)
                    si.unrouteIntraSiteNet(si.getSitePinInst("SR").getBELPin(), si.getBELPin(FF, "SR"));
            si.addSitePIP(si.getSitePIP("SRUSEDMUX", "0"));
        } else {
            if (si.getSitePIPStatus(si.getSitePIP("SRUSEDMUX", "0")) == SitePIPStatus.ON)
                for (String FF : FF_BELS)
                    si.unrouteIntraSiteNet(si.getBELPin("SRUSEDGND", "0"), si.getBELPin(FF, "SR"));
            si.addSitePIP(si.getSitePIP("SRUSEDMUX", "IN"));
        }
    }
    if (CENet != null) {
        if (CENet.isVCCNet()) {
            if (si.getSitePIPStatus(si.getSitePIP("CEUSEDMUX", "IN")) == SitePIPStatus.ON)
                for (String FF : FF_BELS)
                    si.unrouteIntraSiteNet(si.getSitePinInst("CE").getBELPin(), si.getBELPin(FF, "CE"));
            si.addSitePIP(si.getSitePIP("CEUSEDMUX", "1"));
        } else {
            if (si.getSitePIPStatus(si.getSitePIP("CEUSEDMUX", "1")) == SitePIPStatus.ON)
                for (String FF : FF_BELS)
                    si.unrouteIntraSiteNet(si.getBELPin("CEUSEDGND", "1"), si.getBELPin(FF, "CE"));
            si.addSitePIP(si.getSitePIP("CEUSEDMUX", "IN"));
        }
    }
} // end rerouteFFSrCeNets()
\end{lstlisting}
\begin{multicols}{2}

### 2025-05-30 latex/paper/sections/10-placement.tex
site swap
chain swaps
chain swap pseudocode

{
    \centering
    \includegraphics[width=\columnwidth]{figures/placement/moveSingleSite.png}
    \captionof{figure}{Single Site movement}
    \label{fig:moveSingleSite}
}

>>>>

{
    \centering
    \includegraphics[width=\columnwidth]{figures/placement/swapSingleSite.png}
    \captionof{figure}{Single Site Swap Proposal}
    \label{fig:swapSingleSite}
}

\begin{lstlisting}[language=java, caption={Chain Swapping}, label={lst:chain_swap_pseudocode}]
protected void moveSiteChains(List<List<SiteInst>> chains) {
for (List<SiteInst> current_chain : chains) {
    int ch_size = current_chain.size();
    /*

    1) Draw a vertical window of equal size to the current chain. We will refer to this window as the "home window". 
    2) Select a new candidate anchor for this chain.
    2) Project the home window onto the candidate anchor with anchors coinciding. We will refer to this projected window as the "away window".
    3) Check if there are any resident SiteInsts or SiteInst chains present in the away window. Check if the away window breaks any resident SiteInst chains. 
    4)  IF (the away window breaks any chains): 
            extend the window in either direction to 
            fully accommodate all resident chains. 
        ELSE: GOTO evaluateMove(). 
    5) Project the extended away window back onto the original region with the tail of the window coinciding with the tail of the current chain. This projected window becomes the new home window. 
    7) WHILE (the home window breaks any resident chains):
        shift the window upwards by one Site. 
        IF (number of shifts > home window size - ch_size):
            reject the candidate swap.
            CONTINUE onto next chain
    8) evaluateMove(): HPWL Cost evaluation 
    9) evaluateMoveAcceptance(): hill-climb acceptance with global temperature

    */
}
}
\end{lstlisting}

\newpage
\end{multicols}
{
    \centering
    \includegraphics[width=0.9\columnwidth]{figures/placement/swapSiteChain.png}
    \captionof{figure}{Site Chain Swap Proposal}
    \label{fig:swapChainSite}
}
\begin{multicols}{2}

Select a new anchor Site.
Project a window onto the proposed Site
If the window anchor breaks a chain, extend the window in its direction until the anchor coincides with the chain anchor..
If the window tail breaks a chain, extend the window in its direction until the tail coincides with the chain tail.
